On Secure Delegation of SampBQP with a Classical Client

### 2020-03-18 qpip.tex
Make editorial changes on QPIP1 SampBQP

Here we create a $\QPIP_1$ scheme for $\SampBQP$ by constructing a history XZ local Hamiltonian instance, and ask the prover to provide the ground state.
\Ethan{Properly intro and point out the difference between BQP and SampBQP and the challenge here.}
In the case of decision problems, it would be a simple Chernoff bound as in \cite{kitaev2002classical}. A prover, honest or not, couldn't do better than sending $n$ identical copies of some state. The same reasoning does not work in our case, so we give the following construction instead.

First, we make an observation.

@@

\begin{lemma}
	\label{ProtoGroundStateCheck}
	There is a $\QPIP_1$ protocol $\PiGS$ for \cref{AlgGroundStateCheck}. Furthermore, in this protocol, the verifier doesn't need to apply any gates, and only needs to apply $X$ and $Z$ measurements.
\end{lemma}
\begin{prf}
	The prover simply sends the input qubit-by-qubit for the verifier to measure. It is clear that any attacks on this scheme trivially maps onto an attack on the original scheme.
\end{prf}

Now we present \hyperref[ProtoQPIP1]{Protocol~\ref*{ProtoQPIP1}}, a $\QPIP_1$ protocol for any $\SampBQP$ problem with inverse polynomial error in circuit size.

\begin{protocol}{$\QPIP_1$ protocol for $\SampBQP$}
	\label{ProtoQPIP1}
	\begin{itemize}
		\item Let $\varepsilon>0$
		\item Let $m$ be large enough so that $\frac{1}{2^m}$ is negligible in $\varepsilon$
		\item Let $n$ be large enough so both
			$$P\left[Bin(n, \frac{1}{2}-2\varepsilon)\geq\left(\frac{1}{2}-\varepsilon\right)n\right]$$
			and
			$$P\left[Bin(n, \frac{1}{2})\leq\left(\frac{1}{2}-\varepsilon\right)n\right]$$ are negligible in $\varepsilon$ by Chernoff bound
		\item Let $N$ be large enough so that $$\sqrt{\frac{2(m+n)^2\ln\abs{A}}{N-(m+n)}}<\varepsilon$$
			where $\abs{A}$ is the number of qubits in $\ket{\psi_{C(x)}}$
	\end{itemize}
	\begin{enumerate}
		\item The honest prover sends $N$ identical copies of $\ket{\psi_{C'(x)}}$, where $C'$ is $C$ padded with identities as mentioned in \hyperref[idpadding]{Observation~\ref*{idpadding}}
		\item The verifier randomly picks $m$ copies for outputting and $n$ copies for checking
		\item Set counter $c=0$
		\item Set $y=\bot$
		\item For $i$ from $1$ to $N$:
		\begin{enumerate}
			\item If the $i$-th register is for checking:
			\begin{enumerate}
				\item The verifier runs \cref{AlgGroundStateCheck} as described by \hyperref[ProtoGroundStateCheck]{Observation~\ref*{ProtoGroundStateCheck}}.
				\item If the test passes, increment counter $c$.
			\end{enumerate}
			\item If the register is for outputting:
			\begin{enumerate}
				\item The verifier measures the copy in standard basis
				\item If the measurement result of the clock register $\geq\frac{T}{2}$ and $y=\bot$, then update $y$ as the measurement result of the data register.
			\end{enumerate}
			\item Otherwise, just measure the entire register in standard basis and discard it.
		\end{enumerate}
		\item If $\frac{c}{n}\geq\frac{1}{2}-\varepsilon$:
		\begin{enumerate}
			\item If $y\ne\bot$ then output $y$
			\item Otherwise, reject.
		\end{enumerate}
		\item Otherwise, reject.
	\end{enumerate}
\end{protocol}

\begin{thm}
    \label{QPIP1thm}
	For any $L\in\SampBQP$, \hyperref[ProtoQPIP1]{Protocol~\ref*{ProtoQPIP1}} is a protocol for $L$ with negligible completeness. $\forall c\in\bbN$, it achieves soundness $(\varepsilon, \varepsilon)$ where $\varepsilon=T^{-c}$. Furthermore, in this protocol, the verifier doesn't need to apply any gates, and only needs to apply $X$ and $Z$ measurements.
\end{thm}
\begin{prf}
	It clearly has completeness negligibly close to $1$. Now we analyze soundness against a possibly cheating prover that has at least $\delta$ probability to be accepted. Let $\sigma$ be the state the prover sends to the verifier.

	Picking $n+m$ out of $N$ registers is equivalent to first applying a random permutation then taking the first $n+m$ registers.
	A random permutation is equivalent to a classical mix over all possible permutations of $\sigma$:

@@

	So we can apply \cref{deFinetti}. $\exists\set{w_i, \rho_i}$ such that:
	$$\max_{\Lambda_i}\norm{\Lambda_1\otimes\ldots\otimes\Lambda_{m+n}(\sigma'-\sum_i w_i\rho_i^{\otimes n+m})}=O(\varepsilon)$$

@@

	Let $p_i$ be the accept probability of at least $(\frac{1}{2}-\varepsilon)n$ copies from $\rho_i^{\otimes n}$ accept under \cref{AlgGroundStateCheck}.
	So the accept chance of this prover is $\sum_j w_j p_j>\delta$
	By Bayes' theorem, conditioned on acceptance, the probability of $\tilde{\sigma}$ being $\rho_i^{\otimes n+m}$ is then $\frac{w_i p_i}{\sum_j w_j p_j}$.

@@

	The probability of measuring $t\geq T$ on the clock register of $\ket{\psi_{C'(x)}}$ is $\frac{1}{2}$, so the probability of failing to do so after $m$ attempts is negligible.
	Once we do measure $t\geq T$, the data register is $C(x)$.

>>>>

In this section, we construct a one-message $\QPIP_1$ delegation protocol for $\SampBQP$.
It is statistically secure with negligible completeness errors and inverse polynomial soundness errors.
At a high level idea, it is a cut-and-choose protocol.
The server constructs multiple copies of the ground state as certificates,
then the client randomly chooses a copy to output and checks the rest.

There are two main challenges to this cut-and-choose approach.
First, the client needs to reliably extract the circuit output from its corresponding ground state.
We accomplish this by padding the circuit with identity gates at the end.
By doing so, the clock register collapses to after the last non-identity gate with high probability.

The other challenge is that a cheating prover's copies of certificates can be arbitrarily entangled,
so the usual techniques such as Chernoff bounds cannot be applied as-is.
This isn't a challenge for our $\Piblind$ for $\BQP$ earlier because its soundness is one-sided:
we only guarantee that a client cannot accept a no-instance;
a cheating server is allowed to cause the client to reject a yes-instance.
Because of this asymmetry, a cheating prover's for $\Piblind$ optimal strategy is to maximize its acceptance probability,
which corresponds to sending $n$ identical copies of some state.
In our case with $\SampBQP$, we overcome this challenge by using de Finetti's theorem
to approximate our measurement results with that of measurement results on some unentangled copies.

@@

\def\Samp{\mathsf{Samp}}
\nc{\PiSamp}{\ensuremath{\Pi_\Samp}}
\nc{\VSamp}{\ensuremath{V_\Samp}}
\nc{\PSamp}{\ensuremath{P_\Samp}}
\nc{\PSampstar}{\ensuremath{P_\Samp^*}}
\nc{\cVSamp}[1]{\ensuremath{\cV_{\Samp,#1}}}
\nc{\cPSamp}[1]{\ensuremath{\cP_{\Samp,#1}}}

\def\GS{\mathsf{GS}}

Now we present our $\QPIP_1$ protocol. It is parametrized by $\lambda$, the degree of its inverse polynomial soundness.

\begin{protocol}{$\QPIP_1$ protocol $\PiSamp$ for $\SampBQP$}
	\label{ProtoQPIP1}
	Let $\varepsilon=\frac{1}{T^\lambda}$.
	Let $C'$ be $C$ padded with $T^{\lambda + 1}$ identity gates at the end.
	Let $n$ be large enough so by Chernoff bound:
		$$P\left[Bin(n, \frac{1}{2}-2\varepsilon)\geq\left(\frac{1}{2}-\varepsilon\right)n\right]<\varepsilon$$
		$$P\left[Bin(n, \frac{1}{2})\leq\left(\frac{1}{2}-\varepsilon\right)n\right]<\varepsilon$$
	Let $N$ be large enough so that de Finetti's theorem can be applied at the cost of $\varepsilon$ errors if one randomly picks $n+1$ subsystems out of $N$ permutation-invariant subsystems.
		That is,
		$$\sqrt{\frac{2(n+1)^2\ln\abs{A}}{N-(n+1)}}<\varepsilon$$
		where $\abs{A}$ is the number of qubits in $\ket{\psi_{C'(x)}}$
	\begin{enumerate}
		\item The verifier privately samples $I\subset[N]$ s.t. $\abs{I}=n$, and $k\xleftarrow{\$}[N]\setminus I$.
		\item The honest prover prepares $N$ copies of $\ket{\psi_{C'(x)}}$ and sends all of it to the verifier qubit-by-qubit.
		\item For $i$ from $1$ to $N$, the verifier chooses what to do to $\rho_i$ as follows:
		\begin{enumerate}
			\item If $i\in I$, run $\cA_{\GS}(\rho_i)$.
			\item Otherwise, if $i=k$, measure the data register of $\rho_i$ and save the outcome as $y$.
			\item Otherwise, discard $\rho_i$.
		\end{enumerate}
		\item If the proportion of copies accepted by $\cA_{\GS}$ is greater than $\frac{1}{2}-\varepsilon$ then the verifier accepts and outputs $y$. Otherwise, it rejects.
	\end{enumerate}
\end{protocol}

Note that $\VSamp$ only needs to apply $X$ and $Z$ measurements, and is classical otherwise. $\PiSamp$ has negligible completeness errors by construction. We now show its soundness.

\begin{thm}
    \label{QPIP1thm}
	$\PiSamp$ has $O(\frac{1}{T^\lambda})$ soundness error.
\end{thm}
\begin{prf}
	Suppose $\PSampstar$ is a cheating prover that sends some $\sigma$ to the verifier.

	We first show that de Finetti's theorem can be used here.

	Randomly picking $n+1$ out of $N$ registers is equivalent to first applying a random permutation then taking the first $n+1$ registers.
	A random permutation, in turn, is equivalent to a classical mix over all possible permutations of $\sigma$:

@@

	Now we apply \cref{deFinetti}. $\exists\set{w_i, \rho_i}$ such that:
	$$\max_{\Lambda_i}\norm{\Lambda_1\otimes\ldots\otimes\Lambda_{m+n}(\sigma'-\sum_i w_i\rho_i^{\otimes n+m})}=O(\varepsilon)$$

@@

	Let $p_i$ be the accept probability of at least $(\frac{1}{2}-\varepsilon)n$ copies from $\rho_i^{\otimes n}$ accept under \cref{AlgGroundStateCheck}.
	So the accept chance of this prover is $\sum_j w_j p_j>\delta$
	By Bayes' theorem, conditioned on acceptance, the probability of $\tilde{\sigma}$ being $\rho_i^{\otimes n+1}$ is then $\frac{w_i p_i}{\sum_j w_j p_j}$.

@@

	The probability of measuring $t<T$ on the clock register of $\ket{\psi_{C'(x)}}$ is $\frac{T}{T+T^{\lambda+1}}<\frac{1}{T^\lambda}=\varepsilon$,
	so the data register has $1-\varepsilon$ probability to be $C(x)$ at this point.
